---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Finn Liebig (12402643)"
date: "`r Sys.Date()`"
output: distill::distill_article
editor_options: 
  markdown: 
    wrap: 72
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Matrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to its bullet points or use a new level 3 heading if you want.
-   Always render the R code for your solutions (`echo=TRUE`) and make
    sure to include the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown or Quarto document. Submit the files separately on moodle,
    **not** as a zip archive.
-   Submission format is HTML. Other formats will lead to a deduction of
    points.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `2025-exeRcise-sheet-3`. If you forgot how to do this, you can
    follow this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)

    **Vorteile**

-   *Versionskontrolle*: Mithilfe von Git lässt sich die
    Bearbeitungshistorie des Dokuments zurückverfolgen. Damit ist es
    nicht nur möglich Back-up Dateien zu erstellen, sondern es ist
    möglich auf einzelne Arbeitsschritte sprich Änderungen zuzugreifen.
    Das erleichtert u.a. eventuelle Fehlerbehebungen und stellt eine
    weitere Sicherheitsvorkehrung beim Speichern des Dokuments dar.

-   *Reproduzierbarkeit*: Dadurch, dass das Repository geteilt werden
    kann, ist es möglich, dass Dritte die vorgenommenen
    Bearbeitungsschritte in ihrem Projekt leichter nachvollziehen können
    und bei Bedarf auch reproduzieren können. Das fördert den
    wissenschaftlichen Austausch im allgemeinen ebenso wie die Kontrolle
    und Verbesserung bei wissenschaftlicher Arbeiten.

    **Nachteile**

-   *Arbeit mit der Befehlszeile*: Git selbst besitzt kein ordentliches
    IDE. Stattdessen muss ein Großteil der Befehle immer über die
    Befehlszeile eingegeben werden. Das kann umständlich und für
    Anfänger ungeeignet sein. Einige IDEs wie R Studio besitzen
    integrierte IDEs, diese bilden allerdings nur bedingt die Funktionen
    von Git ab.

-   *Leistung*: Besonders große Projekte mit vielen Dateien und
    Datenverzweigungen können bei Git Leistungsprobleme hervorrufen.
    Gerade bei älteren Computern macht sich das dann in der allgemeinen
    Performance der aktiven Programme zu bemerken und kann eventuell zu
    Abstürzen führen. Des weiteren kann Git nicht im Hintergrund laufen,
    was ebenfalls mehr Arbeitsspeicher erfordern kann.

e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves. You
    can, but don't need to create a `.gitignore` file. Just do not
    commit these files manually.

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Library

```{r}
library(remotes)
library(knitr)
library(tidyverse)
library(readr)
library(kableExtra)
library(easystats)
library(ggstats)
library(broom)
library(coefplot)
library(effects)
library(reticulate)
library(nnet)
```

## Exercise 3: Pixar Films (4 Points)

Download the `pixar_films` and `public_response` datasets from the
GitHub repository and track them in git.

Link:
<https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-03-11>

For small datasets like these adding them to git is not a problem.

a)  Load the `pixar_films` dataset into R. Clean the dataset by removing
    films without a title. Inspect the variable `film_rating`. What are
    the possible values and what do they mean? Create a factor variable
    for the film rating. Why is this appropriate?

    ```{r}
    pixar_films <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/pixar_films.csv')
    ```

    ```{r}
    pixar_films_clean <- pixar_films %>% filter(!is.na(film), film != "")
    unique(pixar_films_clean$film_rating)
    ```

    Die möglichen Werte sind G (General Audience: Für alle Altersgruppen
    geeignet) sowie PG (Parental Guidance: Einige Inhalte könnten
    ungeeignet für Kinder sein). Über die Filme mit Wert N/A lässt sich
    keine Aussage treffen.

Ein Faktor ist in R sinnvoll, wenn es sich um kategorische Daten mit
natürlicher Reihenfolge handelt – wie bei den Filmratings "G" und "PG".
Diese spiegeln zunehmende Alterseinschränkungen wider.

Faktoren bewahren diese Reihenfolge, was bei Sortierungen,
Visualisierungen oder statistischen Modellen hilfreich ist. Zudem sparen
sie Speicherplatz und unterstützen eine saubere Datenanalyse.

```{r}
pixar_films_superclean <- pixar_films_clean %>% filter(film_rating != "N/A") %>%
  mutate(film_rating_factor = factor(film_rating, 
                                     levels = c("G", "PG"),
                                     ordered = TRUE))
```

b)  Inspect the film titles manually. Which films form a film series? A
    film series can be identified by a common word in the titles of the
    films, often in conjunction with a number in the title, e.g.
    "Despicable Me" and "Despicable Me 2". Create a dataframe which
    displays a list of the different series with the titles of the films
    and how many films belong to the series. Output the dataframe.

```{r}
pixar_series_df <- data.frame(
  series = c("Toy Story", "The Incredibles", "Cars", "Monsters", "Finding"),
  film_titles = c(
    "Toy Story; Toy Story 2; Toy Story 3; Toy Story 4",
    "The Incredibles; Incredibles 2",
    "Cars; Cars 2; Cars 3",
    "Monsters, Inc.; Monsters University",
    "Finding Nemo; Finding Dory"
  ),
  number_of_films = c(4, 2, 3, 2, 2),
  stringsAsFactors = FALSE
)
print(pixar_series_df)
```

c)  Load the `public_response` dataframe into R. Convert the
    `cinema_score` variable into a factor while ensuring the factor
    levels are defined in ascending order, from the lowest to the
    highest score. Combine `public_response` with the `pixar_films`
    dataset using an appropriate merge variable.

```{r}
public_response <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/public_response.csv')
```

```{r}
public_response_clean <- public_response %>%
  filter(!is.na(cinema_score), cinema_score != "") %>%
  mutate(cinema_score = factor(cinema_score,
                               levels = c("A-", "A", "A+"),
                               ordered = TRUE))
public_response_clean %>% arrange(cinema_score)
```

```{r}
pixar_combined <- left_join(pixar_films_superclean, public_response_clean, 
                            by = c("film" = "film"))
```

d)  Choose one of the variables representing the public response and
    create a bar plot for the films belonging to a series. Here are the
    details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars
        are grouped by film under their respective series on the x-axis.
        Order the bars within a series according to the release date of
        the films.
    -   A title and axis labels for context.

    What do you notice when comparing the scores of the films in a
    series? Do you see any patterns?

```{r}
films_df <- pixar_series_df %>%
  separate_rows(film_titles, sep = "; ") %>%
  rename(title = film_titles)
films_df$rotten_tomatoes <- c(
  100, 100, 98, 97,       # Toy Story 1–4
  97, 93,              # Incredibles 1–2
  75, 40, 69,          # Cars 1–3
  96, 80,              # Monsters Inc. + University
  99, 94               # Finding Nemo + Dory
)
```

```{r}
ggplot(films_df, aes(x = series, y = rotten_tomatoes, fill = title)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Audience Scores of Pixar Film Series",
    x = "Film Series",
    y = "Score on Rotten Tomatoes (%)",
    fill = "Film Title"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Es fällt auf, dass der erste Film einer Reihe immer der höchstbewerteste Film der Reihe ist. Ausnahme ist "Toy Story", bei welcher die ersten beiden Filme gleich hoch gewertet wurden. 

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
b)  Download the dataset and track it in git.
c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: ...
-   Link: <https://github.com/owid/owid-datasets/>...

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R

    Q1: Welche Waffentypen wurden weltweit am häufigsten bei
        Terroranschlägen verwendet und zeigt sich über die Jahre hinweg ein
        Trend in der Nutzung bestimmter Waffenarten?
    
    Q2: Lässt sich anhand der relativen Waffenanteile pro Land voraussagen, ob ein Land in Nordamerika         liegt oder nicht??
    
e)  Use R to answer your chosen question. Please limit your analysis to
    the functions and techniques we have covered so far in the course.
    You are **not expected** to use advanced statistical models or
    external packages which haven't been introduced.

    ```{r}
terrorist_weapon <- read.csv("data/Terrorist attacks by weapon type - Global Terrorism Database (2018).csv")
    ```

    Q1:

```{r}
waffen_spalten <- c(colnames(terrorist_weapon[3:12]))
terrorist_weapon_long <- terrorist_weapon %>%
  pivot_longer(cols = all_of(waffen_spalten), 
               names_to = "Waffentyp", 
               values_to = "Haeufigkeit")
waffen_vis <- terrorist_weapon_long %>%
  group_by(Year, Waffentyp) %>%
  summarise(Haeufigkeit_pro_Jahr = sum(Haeufigkeit, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
ggplot(waffen_vis, aes(x = Year, y = Haeufigkeit_pro_Jahr, color = Waffentyp)) +
  geom_line() +
  labs(title = "Trends der Waffentypen bei Terroranschlägen",
       x = "Jahr",
       y = "Häufigkeit",
       color = "Waffentyp") +
  theme_minimal()
```

    Q2:

    Erstellung der Länderlisten: 
```{r}
asien <- c("Afghanistan", "Armenia", "Azerbaijan", "Bahrain", "Bangladesh", "Bhutan",
           "Brunei", "Cambodia", "China", "Cyprus", "Georgia", "India", "Indonesia",
           "Iran", "Iraq", "Israel", "Japan", "Jordan", "Kazakhstan", "Kuwait",
           "Kyrgyzstan", "Laos", "Lebanon", "Malaysia", "Maldives", "Mongolia",
           "Myanmar", "Nepal", "North Korea", "Oman", "Pakistan", "Palestine",
           "Philippines", "Qatar", "Saudi Arabia", "Singapore", "South Korea",
           "Sri Lanka", "Syria", "Taiwan", "Tajikistan", "Thailand", "Timor-Leste",
           "Turkey", "Turkmenistan", "United Arab Emirates", "Uzbekistan", "Vietnam", "Yemen")
afrika <- c("Algeria", "Angola", "Benin", "Botswana", "Burkina Faso", "Burundi",
            "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros",
            "Congo", "Congo, Democratic Republic of the", "Djibouti", "Egypt", "Equatorial Guinea",
            "Eritrea", "Eswatini", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea",
            "Guinea-Bissau", "Ivory Coast", "Kenya", "Lesotho", "Liberia", "Libya", "Madagascar",
            "Malawi", "Mali", "Mauritania", "Mauritius", "Morocco", "Mozambique", "Namibia",
            "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles",
            "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania",
            "Togo", "Tunisia", "Uganda", "Zambia", "Zimbabwe")
europa <- c("Albania", "Andorra", "Austria", "Belarus", "Belgium", "Bosnia and Herzegovina",
            "Bulgaria", "Croatia", "Czech Republic", "Denmark", "Estonia", "Finland",
            "France", "Germany", "Greece", "Hungary", "Iceland", "Ireland", "Italy",
            "Kosovo", "Latvia", "Liechtenstein", "Lithuania", "Luxembourg", "Malta",
            "Moldova", "Monaco", "Montenegro", "Netherlands", "North Macedonia",
            "Norway", "Poland", "Portugal", "Romania", "Russia", "San Marino",
            "Serbia", "Slovakia", "Slovenia", "Spain", "Sweden", "Switzerland",
            "Ukraine", "United Kingdom", "Vatican City")
nordamerika <- c("Antigua and Barbuda", "Bahamas", "Barbados", "Belize", "Canada",
                 "Costa Rica", "Cuba", "Dominica", "Dominican Republic", "El Salvador",
                 "Grenada", "Guatemala", "Haiti", "Honduras", "Jamaica", "Mexico",
                 "Nicaragua", "Panama", "Saint Kitts and Nevis", "Saint Lucia",
                 "Saint Vincent and the Grenadines", "Trinidad and Tobago", "United States")
suedamerika <- c("Argentina", "Bolivia", "Brazil", "Chile", "Colombia", "Ecuador",
                 "Guyana", "Paraguay", "Peru", "Suriname", "Uruguay", "Venezuela")
ozeanien <- c("Australia", "Fiji", "Kiribati", "Marshall Islands", "Micronesia",
              "Nauru", "New Zealand", "Palau", "Papua New Guinea", "Samoa",
              "Solomon Islands", "Tonga", "Tuvalu", "Vanuatu")
```
    
```{r}
terrorist_weapon_long_q2 <- terrorist_weapon_long %>%
  mutate(Kontinent = case_when(
    Entity %in% afrika ~ "Afrika",
    Entity %in% asien ~ "Asien",
  Entity %in% europa ~ "Europa",
   Entity %in% nordamerika ~ "Nordamerika",
   Entity %in% suedamerika ~ "Südamerika",
    Entity %in% ozeanien ~ "Ozeanien",
    TRUE ~ "Unbekannt"
  ))
waffen_kontinent <- terrorist_weapon_long_q2 %>%
  filter(Kontinent != "Unbekannt") %>%
  group_by(Kontinent, Waffentyp) %>%
  summarise(Gesamt = sum(Haeufigkeit, na.rm = TRUE), .groups = "drop") %>%
  group_by(Kontinent) %>%
  mutate(Anteil = Gesamt / sum(Gesamt)) %>%
  ungroup()
```

Anpassung Datensatz

```{r}
weapon_wide <- terrorist_weapon_long_q2 %>%
  group_by(Entity, Waffentyp) %>%
  summarise(Anzahl = sum(Haeufigkeit, na.rm = TRUE), .groups = "drop") %>%  
  ungroup() %>%
  group_by(Entity) %>%
  mutate(Anteil = Anzahl / sum(Anzahl, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Kontinent = case_when(
    Entity %in% afrika ~ "Afrika",
    Entity %in% asien ~ "Asien",
  Entity %in% europa ~ "Europa",
   Entity %in% nordamerika ~ "Nordamerika",
   Entity %in% suedamerika ~ "Südamerika",
    Entity %in% ozeanien ~ "Ozeanien",
    TRUE ~ "Unbekannt" 
  )) %>%
   select(-Anzahl) %>%  # Entfernt die 'Anzahl'-Spalte
  pivot_wider(
    names_from = Waffentyp,  # Wandelt die Waffentypen in Spalten um
    values_from = Anteil,    # Verwendet den Anteil als Wert
    values_fill = list(Anteil = 0)  # Füllt leere Werte mit 0
  )
weapon_wide$Kontinent <- as.factor(weapon_wide$Kontinent)
weapon_wide <- weapon_wide %>%
  mutate(Kontinent_bin = if_else(Kontinent == "Nordamerika", 1, 0)) %>%
select(-Entity) %>%
  select(-Kontinent)
```

Modell:

```{r}
model <- glm(Kontinent_bin ~ ., data = weapon_wide, family = binomial)
summary(model)
```

Sentence: Anhand der Ergebnisse lässt zeigen, dass die Forschungsfrage mit einem Nein beantworten lässt, da die Beta-Werte für unser Log-Modell statistisch nicht (besonders) signifikant sind.

f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.

```{r fig-waffenkontinent, fig.cap="Abbildung 1: Durchschnittlicher Anteil verschiedener Waffentypen bei terroristischen Anschlägen, gruppiert nach Kontinent."}
ggplot(waffen_kontinent, aes(x = Kontinent, y = Anteil, fill = Waffentyp)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 1) +
  geom_text(aes(label = round(Anteil, 4), y = -0.04),
            position = position_dodge(width = 0.8),
            vjust = 0, size = 2, angle = 90) +
  labs(
    title = "Verwendung von Waffentypen nach Kontinent",
    x = "Kontinent",
    y = "Gesamthäufigkeit",
    fill = "Waffentyp"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  coord_cartesian(ylim = c(-0.05, max(waffen_kontinent$Anteil) * 1.1))
```

## Final Note

Make sure to push all your commits and changes to GitHub before
submitting the exercise sheet.
